Modelling BCG vaccination in the UK

### 2019-12-16 rmds/08-model-development.Rmd
added explanation and check 8.1 -> 8.15 equations

Where $\iota_i$ is the age stratified number of non-UK born pulmonary cases notified in a given year, $P^{u,v}_i$ is the number of vaccinated, and unvaccinated, pulmonary TB cases, $\rho_i$ is the age-specific proportion of cases that are smear positive, $\nu^P_i$ is the age-specific rate of starting treatment for active pulmonary TB, $C_{ai}$ is the age-stratified contact matrix (Section \@ref(contact-matrix)), $\beta_a$ is the age-stratified transmission probability, $N_a$ is the number of people in age group $a$, and $M_i$ is the age-stratified mixing rate between the UK born and non-UK born population.

@@

The probability of transmission can be defined as the probability that a single contact between an infectious active TB case and a susceptible individual will lead to TB infection. The probability of transmission ($\beta_a$) can be redefined in terms of effective contacts (the transmission probability times the contact rate; $C_{\text{eff}}$), historic effective contacts ($C^{\text{hist}}_{\text{eff}}$), actual average yearly total contacts ($C_{\text{actual}}$), the average period of time infectious ($\frac{1}{\nu_{\text{avg}}^{P}}$), and the average mortality rate ($\mu_{\text{avg}}$) as follows,

@@

Vynnycky et al. found that the effective contact rate for TB was approximately 22 in 1900 and fell to approximately 1 in 1990.[@Vynnycky1999] Incidence rates have increased since the early 1980s and it is unclear what impact this has had on the effective contact rate. I have assumed that the effective contact rate is normally distributed with a mean of 1 and a standard deviation of 0.5. For the historic effective contact rate I have assumed a uniform distribution with a lower bound of the current effective contact rate and an upper bound of 20. I have also assumed that the historic contact rate declines over time, starting in 1935, reducing to the current effective contact rate in 1980. The speed of this decay is set by defining a decay parameter ($C^{\text{half-life}}_{\text{eff}}$). The prior for the decay parameter is assumed to be normally distributed with a mean of 5 years and a standard deviation of 5 (truncated to be greater than 0). This prior is based on the observed trend in notifications. Additional age stratification of $\beta$ is explored by including modifiers for certain age-groups. The baseline scenario is that no modification is required, with variation explored for young adults (15-29; $\beta_{\text{young adult}}$) as a scenario. The prior for this modifier was assumed to be uniform, bounded by 0 and 10. The contact rate is estimated by averaging the total age-specific contact rates estimated from POLYMOD data (Section \@ref(contact-matrix)) on an annual basis.

>>>>

Where $\iota_i$ is the age stratified number of non-UK born pulmonary cases notified in a given year, $P^{u,v}_i$ is the number of vaccinated, and unvaccinated, pulmonary TB cases, $\rho_i$ is the age-specific proportion of cases that are smear positive, $\nu^P_i$ is the age-specific rate of starting treatment for active pulmonary TB, $C_{ai}$ is the age-stratified contact matrix (Section \@ref(contact-matrix)), $\beta_a$ is the age-stratified transmission probability, $N_a$ is the number of people in age group $a$, and $M_i$ is the age-stratified mixing rate between the UK born and non-UK born population. Non-UK born cases pulmonary cases ($\iota_i$) are included in the the force of infection, along with their own mixing rate ($M_i$), and the duration of time spent infectious ($\frac{1}{\nu^P_i}$) so that their external forcing on UK born cases is modelled (Section \@ref(model-just)). A separate mixing rate ($M_i$) is used as the extent of this forcing is not known. Both UK born and non-UK born pulmonary TB cases are multiplied by the proportion of cases that are smear positive ($\rho_i$) as it is thought that these cases are responsible for the majority of onwards transmission (Section \@ref(model-just)).

@@

The probability of transmission can be defined as the probability that a single contact between an infectious active TB case and a susceptible individual will lead to TB infection. The probability of transmission ($\beta_a$) can be redefined in terms of effective contacts (the number of contacts that are infected from each infected case; $C_{\text{eff}}$), historic effective contacts ($C^{\text{hist}}_{\text{eff}}$), actual average yearly total contacts ($C_{\text{actual}}$), the average period of time infectious ($\frac{1}{\nu_{\text{avg}}^{P}}$), and the average mortality rate ($\mu_{\text{avg}}$) as follows,

@@

Vynnycky et al. found that the effective contact rate for TB was approximately 22 in 1900 and fell to approximately 1 in 1990.[@Vynnycky1999] Incidence rates have increased since the early 1980s and it is unclear what impact this has had on the effective contact rate. I have assumed that the effective contact rate is normally distributed with a mean of 1 and a standard deviation of 0.5. For the historic effective contact rate I have assumed a uniform distribution with a lower bound of the current effective contact rate and an upper bound of 20. I have also assumed that the historic contact rate declines over time, starting in 1935, reducing to the current effective contact rate in 1980. The speed of this decay is set by defining a decay parameter ($C^{\text{half-life}}_{\text{eff}}$). The prior for the decay parameter is assumed to be normally distributed with a mean of 5 years and a standard deviation of 5 (truncated to be greater than 0). This prior is based on the observed trend in notifications. Age stratification of $\beta_a$ is explored by including modifiers for certain age-groups. The baseline scenario is that no modification is required, with variation explored for young adults (15-29; $\beta_{\text{young adult}}$) as a scenario. The prior for this modifier was assumed to be uniform, bounded by 0 and 10. The contact rate is estimated by averaging the total age-specific contact rates estimated from POLYMOD data (Section \@ref(contact-matrix)) on an annual basis.
